Evolution of pre-copulatory and post-copulatory strategies of inbreeding avoidance

### 2017-07-19 ms_pre_post.Rmd
Change abstract

---
title: Evolution of pre-copulatory and post-copulatory strategies of inbreeding avoidance
  and associated polyandry
author: ''
date: "4 May 2017"
output:
  word_document:
    fig_caption: yes
    pandoc_args:
    - --csl
    - evolution.csl
    reference_docx: docx_template.docx
  pdf_document:
    fig_caption: yes
  html_document: default
header-includes:
- \usepackage{natbib}
- \usepackage{lineno}
- \linenumbers
- \bibliographystyle{amnatnat}
linestretch: 2
csl: evolution.csl
bibliography: ms_pre_post.bib
---

@@

Inbreeding depression is widely hypothesised to drive adaptive evolution of pre-copulatory and post-copulatory mechanisms of inbreeding avoidance, which in turn are hypothesised to affect evolution of polyandry (i.e., female multiple mating). However, surprisingly little theory or modelling critically examines selection for pre-copulatory or post-copulatory inbreeding avoidance, or both strategies, given evolutionary constraints and direct costs, or examines how evolution of inbreeding avoidance strategies might feed back to affect evolution of polyandry. Selection for post-copulatory inbreeding avoidance, but not for pre-copulatory inbreeding avoidance, requires polyandry, while interactions between pre-copulatory and post-copulatory inbreeding avoidance might cause functional redundancy (i.e., 'degeneracy') potentially generating complex evolutionary dynamics among inbreeding strategies and polyandry. We use individual-based modelling to quantify evolution of interacting pre-copulatory and post-copulatory inbreeding avoidance and associated polyandry given strong inbreeding depression and different evolutionary constraints and direct costs. We find that evolution of post-copulatory inbreeding avoidance increases selection for initially rare polyandry, and that selection for a costly inbreeding avoidance strategy was weak or negligible given a cost free alternative strategy. Further, fixed pre-copulatory inbreeding avoidance often completely precluded evolution of polyandry and hence post-copulatory inbreeding avoidance, but fixed post-copulatory inbreeding avoidance did not preclude evolution of pre-copulatory inbreeding avoidance. Evolution of inbreeding avoidance phenotypes and associated polyandry are therefore affected by evolutionary feedbacks and degeneracy. All else being equal, evolution of pre-copulatory inbreeding avoidance and resulting low polyandry is more likely when post-copulatory inbreeding avoidance is precluded or costly, and evolution of post-copulatory inbreeding avoidance greatly facilitates evolution of costly polyandry.

>>>>

---
title: Evolution of pre-copulatory and post-copulatory strategies of inbreeding avoidance
  and associated polyandry
author: ''
date: "4 May 2017"
output:
  pdf_document:
    fig_caption: yes
  html_document: default
  word_document:
    fig_caption: yes
    pandoc_args:
    - --csl
    - evolution.csl
    reference_docx: docx_template.docx
header-includes:
- \usepackage{natbib}
- \usepackage{lineno}
- \linenumbers
- \bibliographystyle{amnatnat}
linestretch: 2
csl: evolution.csl
bibliography: ms_pre_post.bib
---

@@

Inbreeding depression is widely hypothesised to drive adaptive evolution of pre-copulatory and post-copulatory mechanisms of inbreeding avoidance, which in turn are hypothesised to affect evolution of polyandry (i.e., female multiple mating). However, surprisingly little theory or modelling critically examines selection for pre-copulatory or post-copulatory inbreeding avoidance, or both strategies, given evolutionary constraints and direct costs, or examines how evolution of inbreeding avoidance strategies might feed back to affect evolution of polyandry. Selection for post-copulatory inbreeding avoidance, but not for pre-copulatory inbreeding avoidance, requires polyandry, while interactions between pre-copulatory and post-copulatory inbreeding avoidance might cause functional redundancy (i.e., 'degeneracy') potentially generating complex evolutionary dynamics among inbreeding strategies and polyandry. We use individual-based modelling to quantify evolution of interacting pre-copulatory and post-copulatory inbreeding avoidance and associated polyandry given strong inbreeding depression and different evolutionary constraints and direct costs. We find that evolution of post-copulatory inbreeding avoidance increases selection for initially rare polyandry, and that selection for a costly inbreeding avoidance strategy was weak or negligible given a lower cost alternative strategy. Further, fixed pre-copulatory inbreeding avoidance often completely precluded evolution of polyandry and hence post-copulatory inbreeding avoidance, but fixed post-copulatory inbreeding avoidance did not preclude evolution of pre-copulatory inbreeding avoidance. Evolution of inbreeding avoidance phenotypes and associated polyandry are therefore affected by evolutionary feedbacks and degeneracy. All else being equal, evolution of pre-copulatory inbreeding avoidance and resulting low polyandry is more likely when post-copulatory inbreeding avoidance is precluded or costly, and evolution of post-copulatory inbreeding avoidance greatly facilitates evolution of costly polyandry.
